---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# stac4cast

<!-- badges: start -->
[![R-CMD-check](https://github.com/eco4cast/stac4cast/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/eco4cast/stac4cast/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The goal of stac4cast is to generate [STAC](https://stacspec.org/en) catalogs for forecasts.

## Installation

You can install the development version of stac4cast from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("eco4cast/stac4cast")
```

## Overview

The STAC metadata standard essentially defines three levels of objects:

- Items
- Catalogs
- Collections

An item is a core atomic unit of data. A catalog is a list of items, a collection is an extension of a catalog that adds additional metadata. These objects are represented as JSON files on a static website, linked by URLs. Once created, STAC can be visualized and explored using tools such as [stac browser](https://radiantearth.github.io/stac-browser/#/) and queried programmatically with packages such as [rstac](https://cran.r-project.org/web/packages/rstac)

## Getting started


```{r ex}
library(stac4cast)
```

See the [examples directory](https://github.com/eco4cast/stac4cast/tree/main/inst/examples) for current examples. 
